### 2019-04-03 Thesis_Latex/ThesisFinal/Introduction/Introduction.tex
update final thesis

\chapter{Introduction}
%\addcontentsline{toc}{chapter}{Introduction}
In robotic research, the problem of navigation is among the most important. Basically, all autonomous mobile robots need some kind of navigation abilities to perform, localization, motion planning and guidance \cite{Skoda:Thesis:2016}. In the present context, we focus on navigation as a process of planning a path of a mobile robot from its current position to a desired goal location, following the planned path, and avoiding any discovered obstacles along the way. The desired paths have to fulfill several conditions to ensure safety and feasibility of the navigation. Moreover, the paths can be also defined in terms of specifications; for example, short or smooth paths are usually more desirable than long and curved ones in every dynamic environments.

Beyond the path planning, the navigation problem also involves reacting to changes of the environment model. Robots are required to move towards the target in a short time and avoid either static or dynamic obstacles observed by their sensors, which involves efficient path planning and valid obstacle avoidance.

Research and development of Unmanned Ground Vehicles has been dominated by DARPA (Defense Advanced Research Projects Agency) and NASA (National Aeronautics and Space Administration). The DARPA initiative started with the development of the first mobile robot, Shakey, and also includes the Autonomous Land Vehicle and the DARPA Demo I1 Program. NASA sponsors the development of unmanned vehicles for planetary surface exploration, from the Jet Propulsion Laboratory Mars Rover to the most recent Mars Pathfinder. Recent UGV design and development has been enhanced to build UGVs capable of operating in Intelligent Vehicle Highway Systems \cite{Bai2015}.

Over the last decades, the development of Advanced Driver Assistance Systems has become a critical endeavor to attain different objectives: safety enhancement, mobility improvement, energy optimization and comfort \cite{Gruyer2017}. Much of the argument used in this discussion is based on the road mortality we see today. According to data from the World Health Organization, in 2013 there were about 1.25 million road deaths worldwide, and this number is expected to increase in the next decade \cite{world2015global}. Algorithms for autonomous navigation are increasingly robust and reliable and are starting to handle complex situations and decision problems.

According to Katrakazas \cite{Katrakazas2015}, local navigation is responsible for guiding the vehicle, that is, for the planning of the direction and speed to be taken, in a space close to the current position based on information exclusively obtained by the sensors on board. The guiding must be planned in such a way as to guarantee the displacement, from the current state to the objective, without collisions.

The algorithms we have developed at the LAR, follow a new and different approach for an advanced control strategy for autonomous navigation. The idea is that these methods do not replace the algorithms developed previously but are a valid alternative, so that depending on the situation, the vehicle can choose the best strategy to overcome obstacles or solve problems that can occur on the road that need a decision in real time.  Simulation results demonstrate and verify the feasibility and the usefulness of methods considering different scenarios.

@@

\subsection{ATLAS platforms}\label{sec:ATLASplatform}	
The first developed robot (Figure \ref{fig:modelosatlas1}) was based on an aluminum and wood structure. In this prototype only one camera was installed that pointed to a mirror to allow the complete visualization of the road in which the robot circulated. The traction movement was assured by a mechanical differential coupled to the rear wheels and the steering movement was given by a single front wheel. In order to create a model more similar to an ordinary car, the ATLAS group developed the ATLAS 2000 (Figure \ref{fig:modelosatlas2}) in scale (1:4), with which it managed to win the first autonomous driving competition of the FNR in 2006. After several improvements made in ATLAS 2000, in 2008 a new platform, ATLAS MV (Figure \ref{fig:modelosatlas3}) was created. This robot was designed on a smaller scale (1:5), with the intention of being lighter and faster. On board were installed a new steering system, hydraulic braking and an active perception unit. This robot allowed the conquest of new victories in the autonomus driving competitions.
\begin{figure}[!h]
	\centering
		\begin{minipage}[t]{0.32\textwidth}
			\includegraphics[width=\textwidth]{../figure/modelosatlas1.pdf}
			\subcaption{First ATLAS prototype.}
			\label{fig:modelosatlas1}
		\end{minipage}
		\begin{minipage}[t]{0.32\textwidth}
			\includegraphics[width=\textwidth]{../figure/modelosatlas2.pdf}
			\subcaption{ATLAS 2000.}
			\label{fig:modelosatlas2}
		\end{minipage}
		\begin{minipage}[t]{0.32\textwidth}
			\includegraphics[width=\textwidth]{../figure/modelosatlas3.pdf}
			\subcaption{ATLAS MV.}
			\label{fig:modelosatlas3}
		\end{minipage}
		\caption{Some of the ATLAS project small scale platforms (adapted from \cite{Pereira2012}).}
		\label{fig:modelosatlas}
\end{figure}

\subsection{ATLASCAR}\label{sec:ATLASCAR}
Driven by the positive results achieved with scale models and years of navigation experience in controlled environments, in 2010 the Group of Automation and Robotics decided to invest in a large-scale project, ATLASCAR (Figure \ref{fig:atlascar1}). The vehicle used for this project was a Ford Escort Station Wagon of 1998 powered by a gasoline internal combustion engine, in which several sensors, processing units and actuators were installed. On-board sensors processed data collected from the vehicle and its surroundings, with different LIDARs for obstacle detection and environmental reconstruction, pedestrian detection cameras and a Global Navigation Satellite System (GNSS) for location and route planning. After passing through the processing units, these data were sent to the actuators that allowed the movement and execution of the maneuvers in a completely autonomous way on the part of the vehicle. To power all the equipment, a Uninterruptible Power Supply (UPS) was used, loaded from an auxiliary alternator. During this project, many works were developed in the Laboratory for Automation and Robotics, many of which produced master thesis. For example, in 2014 Cabral de Azevedo \cite{Azevedo2014} developed a module to detect pedestrians using sensory fusion of LIDAR and vision data while in 2016 Vieira da Silva \cite{Silva2016} created a multisensory calibration module that was exported to subsequent projects.
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{../figure/atlascar1.jpg}
	\caption{The car used in ATLASCAR, based on Ford Escort Station Wagon of 1998 (adapted from \cite{Pereira2012}).}
	\label{fig:atlascar1}
\end{figure}

\subsection{ATLASCAR2}\label{sec:ATLASCAR2}
Given the different limits, to continue the project, in 2016, a new vehicle was acquired: ATLASCAR2 (Figure \ref{fig:atlascar2}). This time it was chosen as a platform
an electric car, a Mitsubishi iMiEV, with an autonomy range of \SI{100}{km}. The fact that the vehicle is electric allows to use the energy stored in the batteries, making it easier to power the sensors installed. In fact, despite the short time of existence, 3 LIDARs, a camera, inclinometry sensors and a GNSS unit are already installed on the ATLASCAR2. Many of these sensors were transferred from ATLASCAR to this project during the work of Madureira Correia \cite{Madureira2017} in 2017, where a module for detecting free space around the car was also developed while in 2018 Ricardo Silva \cite{Ricardo:Thesis:2018} created a local navigation module for driver assistance in the immediate decision making, identifing a solution based on a multiple hypothesis approach.
\begin{figure}[!h]
	\centering
	\includegraphics[width=\textwidth]{../figure/atlascar2.jpg}
	\caption{The vehicle used in ATLASCAR2, based on an electric car, a Mitsubishi iMiEV of 2015 (adapted from \cite{Ricardo:Thesis:2018}).}
	\label{fig:atlascar2}
\end{figure}

\section{Autonomous Cars}\label{sec:autonomous_examples}
The legal definition of autonomous vehicle in the District of Columbia code is:
\begin{quotation}
	\itshape ''Autonomous vehicle'' means a vehicle capable of navigating District roadways and interpreting traffic-control devices without a driver actively operating any of the vehicle's control systems. The term ''autonomous vehicle'' excludes a motor vehicle enabled with active safety systems or driver-assistance systems, including systems to provide electronic blind-spot assistance, crash avoidance, emergency braking, parking assistance, adaptive cruise control (ACC), lane-keep assistance (LKA), lane-departure warning, or traffic-jam and queuing assistance, unless the system alone or in combination with other systems enables the vehicle on which the technology is installed to drive without active control or monitoring by a human operator. 
\end{quotation}

The modern automobile companies keep coming up with newer autonomous features in their recent models. Technological advancements seen every day in areas like information technology, communication, data analysis and storage etc. is not exclusive to these areas alone. The realm of autonomous cars is also progressing at a rapid rate these days \cite{Bimbraw2015}. Google's development of self-driving technology began in January 2009. The initial objective of the project was to develop a car able to navigate on highways with minimal human intervention. In December 2016, the unit was renamed Waymo; this name derived from its mission, "a new way forward in mobility". Waymo moved to further test its cars on public roads after becoming its own subsidiary.
\begin{figure}[!h]
	\centering
	\begin{minipage}[t]{0.65\textwidth}
		\includegraphics[width=\textwidth]{../figure/veiculos0.png}
		\subcaption{Google's Firefly self-driving prototype in 2015 (adapted from \cite{waymoh}).}
		\label{fig:veiculos0}
	\end{minipage}
	\begin{minipage}[t]{0.65\textwidth}
		\includegraphics[width=\textwidth]{../figure/veiculos1.png}
		\subcaption{Waymo Chrysler Pacifica Hybrid self-driving prototype in 2017 (adapted from \cite{waymoh}).}
		\label{fig:veiculos1}
	\end{minipage}
	\caption{Some of the Waymo/Google prototypes tested across multiple locations in the Unided States in recent years.}
	\label{fig:waymo}
\end{figure}

@@

With the advances in autonomous technology, VIAC or VisLab Intercontinental Autonomous Challenge was one of the major competitions which led to improvements in the testing and analysis of autonomous vehicles and robotics. It was a 13,000 kilometers trip, nearly three months from Parma, Italy to Shanghai, China from July 20, 2010 to October 28, 2010. It involved four autonomous vehicles with negligible human intervention and high level of autonomy \cite{Broggi2010}. One of VisLab's advanced autonomous car, BRAiVE (Figure \ref{fig:veiculos4}), drove in downtown Parma on July 12th, 2013. It successfully navigated narrow rural roads, crosswalks, traffic lights, pedestrian areas, roundabouts and artificial hazards. It was a pioneer in the field of vehicular robotics, since it was totally autonomous \cite{Broggi2013}.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.65\textwidth]{../figure/veiculos4.png}
	\caption{BRAiVE prototype developted by VisLab, based on a Hyundai Sonata (adapted from \cite{Broggi2013}).}
	\label{fig:veiculos4}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.6\textwidth]{../figure/veiculos5.png}
	\caption{Navya Shuttle developted in 2016 by Navya Group in France.}
	\label{fig:veiculos5}
\end{figure}

>>>>

\begin{savequote}[75mm]
	``Self-driving cars are the natural extension of active safety and obviously something we thing we should do."
	\qauthor{Elon Musk}
\end{savequote}

\chapter{Introduction}
%\addcontentsline{toc}{chapter}{Introduction}
In robotic research, the problem of navigation is among the most important. Basically, all autonomous mobile robots need some kind of navigation abilities to perform, localization, motion planning and guidance \cite{Skoda:Thesis:2016}. In the present context, we focus on navigation as a process of planning a path of a mobile robot from its current position to a desired goal location, following the planned path, and avoiding any discovered obstacles along the way. The desired paths have to fulfill several conditions to ensure safety and feasibility of the navigation. Moreover, the paths can be also defined in terms of specifications; for example, short or smooth paths are usually more desirable than long and curved ones in every dynamic environments. Beyond the path planning, the navigation problem also involves reacting to changes of the environment model. Robots are required to move towards the target in a short time and avoid either static or dynamic obstacles observed by their sensors, which involves efficient path planning and valid obstacle avoidance. Research and development of Unmanned Ground Vehicles has been dominated by DARPA (Defense Advanced Research Projects Agency) and NASA (National Aeronautics and Space Administration). The DARPA initiative started with the development of the first mobile robot, Shakey, and also includes the Autonomous Land Vehicle and the DARPA Demo I1 Program. NASA sponsors the development of unmanned vehicles for planetary surface exploration, from the Jet Propulsion Laboratory Mars Rover to the most recent Mars Pathfinder. Recent UGV design and development has been enhanced to build UGVs capable of operating in Intelligent Vehicle Highway Systems \cite{Bai2015}. Over the last decades, the development of Advanced Driver Assistance Systems has become a critical endeavor to attain different objectives: safety enhancement, mobility improvement, energy optimization and comfort \cite{Gruyer2017}. Much of the argument used in this discussion is based on the road mortality we see today. According to data from the World Health Organization, in 2013 there were about 1.25 million road deaths worldwide, and this number is expected to increase in the next decade \cite{world2015global}. Algorithms for autonomous navigation are increasingly robust and reliable and are starting to handle complex situations and decision problems. According to Katrakazas \cite{Katrakazas2015}, local navigation is responsible for guiding the vehicle, that is, for the planning of the direction and speed to be taken, in a space close to the current position based on information exclusively obtained by the sensors on board. The guiding must be planned in such a way as to guarantee the displacement, from the current state to the objective, without collisions. The algorithms we have developed at the LAR, follow a new and different approach for an advanced control strategy for autonomous navigation. The idea is that these methods do not replace the algorithms developed previously but are a valid alternative, so that depending on the situation, the vehicle can choose the best strategy to overcome obstacles or solve problems that can occur on the road that need a decision in real time.  Simulation results demonstrate and verify the feasibility and the usefulness of methods considering different scenarios.

@@

\subsection{ATLAS platforms}\label{sec:ATLASplatform}	
The first developed robot (Figure \ref{fig:modelosatlas1}) was based on an aluminum and wood structure. In this prototype only one camera was installed that pointed to a mirror to allow the complete visualization of the road in which the robot circulated. The traction movement was assured by a mechanical differential coupled to the rear wheels and the steering movement was given by a single front wheel. In order to create a model more similar to an ordinary car, the ATLAS group developed the ATLAS 2000 (Figure \ref{fig:modelosatlas2}) in scale (1:4), with which it managed to win the first autonomous driving competition of the FNR in 2006. After several improvements made in ATLAS 2000, in 2008 a new platform, ATLAS MV (Figure \ref{fig:modelosatlas3}) was created. This robot was designed on a smaller scale (1:5), with the intention of being lighter and faster. On board were installed a new steering system, hydraulic braking and an active perception unit. This robot allowed the conquest of new victories in the autonomus driving competitions.
\begin{figure}[!h]
	\centering
		\begin{minipage}[t]{0.32\textwidth}
			\includegraphics[width=\textwidth]{./figure/modelosatlas1.pdf}
			\subcaption{First ATLAS prototype.}
			\label{fig:modelosatlas1}
		\end{minipage}
		\begin{minipage}[t]{0.32\textwidth}
			\includegraphics[width=\textwidth]{./figure/modelosatlas2.pdf}
			\subcaption{ATLAS 2000.}
			\label{fig:modelosatlas2}
		\end{minipage}
		\begin{minipage}[t]{0.32\textwidth}
			\includegraphics[width=\textwidth]{./figure/modelosatlas3.pdf}
			\subcaption{ATLAS MV.}
			\label{fig:modelosatlas3}
		\end{minipage}
		\caption{Some of the ATLAS project small scale platforms (adapted from \cite{Pereira2012}).}
		\label{fig:modelosatlas}
\end{figure}

\subsection{ATLASCAR}\label{sec:ATLASCAR}
Driven by the positive results achieved with scale models and years of navigation experience in controlled environments, in 2010 the Group of Automation and Robotics decided to invest in a large-scale project, ATLASCAR (Figure \ref{fig:atlascar1}). The vehicle used for this project was a Ford Escort Station Wagon of 1998 powered by a gasoline internal combustion engine, in which several sensors, processing units and actuators were installed. On-board sensors processed data collected from the vehicle and its surroundings, with different LIDARs for obstacle detection and environmental reconstruction, pedestrian detection cameras and a Global Navigation Satellite System (GNSS) for location and route planning. After passing through the processing units, these data were sent to the actuators that allowed the movement and execution of the maneuvers in a completely autonomous way on the part of the vehicle. To power all the equipment, a Uninterruptible Power Supply (UPS) was used, loaded from an auxiliary alternator. During this project, many works were developed in the Laboratory for Automation and Robotics, many of which produced master thesis. For example, in 2014 Cabral de Azevedo \cite{Azevedo2014} developed a module to detect pedestrians using sensory fusion of LIDAR and vision data while in 2016 Vieira da Silva \cite{Silva2016} created a multisensory calibration module that was exported to subsequent projects.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.9\textwidth]{./figure/atlascar1.jpg}
	\caption{The car used in ATLASCAR, based on Ford Escort Station Wagon of 1998 (adapted from \cite{Pereira2012}).}
	\label{fig:atlascar1}
\end{figure}

\subsection{ATLASCAR2}\label{sec:ATLASCAR2}
Given the different limits, to continue the project, in 2016, a new vehicle was acquired: ATLASCAR2 (Figure \ref{fig:atlascar2}). This time it was chosen as a platform
an electric car, a Mitsubishi iMiEV, with an autonomy range of \SI{100}{km}. The fact that the vehicle is electric allows to use the energy stored in the batteries, making it easier to power the sensors installed. In fact, despite the short time of existence, 3 LIDARs, a camera, inclinometry sensors and a GNSS unit are already installed on the ATLASCAR2. Many of these sensors were transferred from ATLASCAR to this project during the work of Madureira Correia \cite{Madureira2017} in 2017, where a module for detecting free space around the car was also developed while in 2018 Ricardo Silva \cite{Ricardo:Thesis:2018} created a local navigation module for driver assistance in the immediate decision making, identifing a solution based on a multiple hypothesis approach.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.85\textwidth]{./figure/atlascar2.jpg}
	\caption{The vehicle used in ATLASCAR2, based on an electric car, a Mitsubishi iMiEV of 2015 (adapted from \cite{Ricardo:Thesis:2018}).}
	\label{fig:atlascar2}
\end{figure}

\section{Autonomous Cars}\label{sec:autonomous_examples}
The legal definition of autonomous vehicle in the District of Columbia code \cite{DColumbia} is:
\begin{quotation}
	\itshape ''Autonomous vehicle'' means a vehicle capable of navigating District roadways and interpreting traffic-control devices without a driver actively operating any of the vehicle's control systems. The term ''autonomous vehicle'' excludes a motor vehicle enabled with active safety systems or driver-assistance systems, including systems to provide electronic blind-spot assistance, crash avoidance, emergency braking, parking assistance, adaptive cruise control (ACC), lane-keep assistance (LKA), lane-departure warning, or traffic-jam and queuing assistance, unless the system alone or in combination with other systems enables the vehicle on which the technology is installed to drive without active control or monitoring by a human operator. 
\end{quotation}

The modern automobile companies keep coming up with newer autonomous features in their recent models. Technological advancements seen every day in areas like information technology, communication, data analysis and storage etc. is not exclusive to these areas alone. The realm of autonomous cars is also progressing at a rapid rate these days \cite{Bimbraw2015}. Google's development of self-driving technology began in January 2009. The initial objective of the project was to develop a car able to navigate on highways with minimal human intervention. In December 2016, the unit was renamed Waymo; this name derived from its mission, "a new way forward in mobility". Waymo moved to further test its cars on public roads after becoming its own subsidiary.
\begin{figure}[!h]
	\centering
	\begin{minipage}[t]{0.45\textwidth}
		\includegraphics[width=\textwidth]{./figure/veiculos0.png}
		\subcaption{Google's Firefly self-driving prototype in 2015 (adapted from \cite{waymoh}).}
		\label{fig:veiculos0}
	\end{minipage}
	\begin{minipage}[t]{0.45\textwidth}
		\includegraphics[width=\textwidth]{./figure/veiculos1.png}
		\subcaption{Waymo Chrysler Pacifica Hybrid self-driving prototype in 2017 (adapted from \cite{waymoh}).}
		\label{fig:veiculos1}
	\end{minipage}
	\caption{Some of the Waymo/Google prototypes tested across multiple locations in the Unided States in recent years.}
	\label{fig:waymo}
\end{figure}

@@

With the advances in autonomous technology, VIAC or VisLab Intercontinental Autonomous Challenge was one of the major competitions which led to improvements in the testing and analysis of autonomous vehicles and robotics. It was a 13,000 kilometers trip, nearly three months from Parma, Italy to Shanghai, China from July 20, 2010 to October 28, 2010. It involved four autonomous vehicles with negligible human intervention and high level of autonomy \cite{Broggi2010}. One of VisLab's advanced autonomous car, BRAiVE (Figure \ref{fig:veiculos4}), drove in downtown Parma on July 12th, 2013. It successfully navigated narrow rural roads, crosswalks, traffic lights, pedestrian areas, roundabouts and artificial hazards. It was a pioneer in the field of vehicular robotics, since it was totally autonomous \cite{Broggi2013}.
\begin{figure}[!h]
	\centering
	\includegraphics[width=0.65\textwidth]{./figure/veiculos4.png}
	\caption{BRAiVE prototype developted by VisLab, based on a Hyundai Sonata (adapted from \cite{Broggi2013}).}
	\label{fig:veiculos4}
\end{figure}

@@

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.6\textwidth]{./figure/veiculos5.png}
	\caption{Navya Shuttle developted in 2016 by Navya Group in France.}
	\label{fig:veiculos5}
\end{figure}

### 2019-04-03 Thesis_Latex/ThesisFinal/frontmatter/abstract.tex
update final thesis

\chapter*{Abstract}
\thispagestyle{empty}
Inserted in the ATLASCAR2 project this work aims to develop a short-term path planning algorithm
for driver assistance in dynamic environments. In order to achieve this objective, it was made a preliminary study of the existing local path planning methods and the projects that have already been developed in this field, their advantages and disadvantages were weighed
and the most successful approaches applied to local navigation in real autonomous driving projects were taken into account.
This thesis presents two different strategies for a self-driving car short-term path planning among multiple moving obstacles. The main task is to study and implement a motion planning and execution framework in order to make ATLASCAR2 coexist with other moving obstacle vehicles by avoiding collision and overtake them when necessary and possible. The first method developted, is an obstacle avoidance system that moves the vehicle around different moving obstacles while the second algorithm is a lane following system that keeps the ATLASCAR2 traveling along the centerline of the lanes on the road. The proposed techniques, based on the adaptive Model Predictive Control paradigm, solve optimization problems formulated in terms of cost minimization under constraints. Simulation results, developted in a MATLAB/Simulink enviroment, demonstrate and verify the feasibility and the usefulness of methods considering different scenarios, opening space for real scenario implementation.

>>>>

%!TEX root = ../dissertation.tex
% the abstract
Inserted in the ATLASCAR2 project this work aims to develop a short-term path planning framework for driver assistance in dynamic environments. In order to achieve this objective, it was made a preliminary study of the existing local path planning methods and the projects that have already been developed in this field, their advantages and disadvantages were weighed and the most successful approaches applied to local navigation in real autonomous driving projects were taken into account.
This thesis presents two different strategies for a self-driving car short-term path planning among multiple moving obstacles. The main task is to study and implement a motion planning and execution framework in order to make ATLASCAR2 coexist with other moving obstacle vehicles by avoiding collision and overtake them when necessary and possible. The first method developted, is an obstacle avoidance system that moves the vehicle around different moving obstacles while the second algorithm is a lane following system that keeps the ATLASCAR2 traveling along the centerline of the lanes on the road. The proposed techniques, based on the adaptive Model Predictive Control paradigm, solve optimization problems formulated in terms of cost minimization under constraints. Simulation results, developted in a MATLAB/Simulink enviroment, demonstrate and verify the feasibility and the usefulness of methods considering different scenarios, opening space for real scenario implementation.
